XOr-Some Bayesian Power analysis

### 2021-09-27 analysis/power/xor-some_bayesian-power.Rmd
Bayesian Power: minimal version

---
title: "XOr-Some Bayesian Power analysis"
author: "Polina Tsvilodub"
date: "9/14/2021"
output: github_document
---

For the main xor-some experiment, we conduct a simulation-based Bayesian power analysis (used in a (loose) sense of determining the required sample size here). This power analysis aims at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least .8. The simulations are based on an assumed effect size $\beta = ±0.15$ for all predictors, and a region of practical equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data. The simulations are further based on the maximal desired model justified theoretically and by the design: $rating = trigger * prior * competence * relevance + REs$

@@

The target response variable is simulated via $\mathcal{N}(\mu_t, 1)$, with $mu_t = \beta_0 + \beta_{pri} + \beta_{comp} + \beta_{rel} + \beta_{trigger} + interactions$. The coefficients for interactions and the trigger effect are assumed to be 0.

@@

# simulate toy data set for one subject 
d_init <-
  # create the 8 conditions
  tibble(prior = rep(c(0, 1), each = 4),
         comp = rep(c(0, 0, 1, 1), times = 2),
         rel = rep(c(0,1), times = 4),
         main_type = as.factor(sample(rep(c("xor", "some"), times=4)))) %>% 
  rowwise() %>%
  # sample effects based on condition
  mutate(
    subj_ID = 1, 
    intercept = 0,
    prior_obs = ifelse(prior == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, -0.15, 0.05)),
    comp_obs = ifelse(comp == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, 0.15, 0.05)),
    rel_obs = ifelse(rel == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, 0.15, 0.05)),
    # sample response
    # what do we do about the interactions? can they safely be ignored?
    target = rnorm(1, 
                   mean = (intercept + prior_obs + comp_obs + rel_obs), 
                   sd = 1) # how to set / determine this properly?
  )
d_init
```

@@

>>>>

---
title: "XOr-Some Bayesian Power analysis"
author: "Polina Tsvilodub"
date: "9/14/2021"
output: 
  github_document:
  pandoc_args: --webtex
---

For the main xor-some experiment, we conduct a simulation-based Bayesian power analysis (used in a (loose) sense of determining the required sample size here). This power analysis aims at determining how many participants are required in order to detect a theoretically motivated conjunctive effect of the predictors prior, relevance and competence with a confidence of at least .8. The simulations are based on an assumed effect size $\beta = ±0.15$ for all predictors, and a region of practical equivalence (ROPE) $\delta = 0.05$ for judging evidence for the directionality of a coefficient given the data. The simulations are further based on the maximal desired model justified theoretically and by the design: `rating = trigger * prior * competence * relevance + REs`

@@

The target response variable is simulated via $\mathcal{N}(\mu_t, 1)$, with $\mu_t = \beta_0 + \beta_\text{pri} + \beta_\text{comp} + \beta_\text{rel} + \beta_\text{trigger} + \text{interactions}$. The coefficients for interactions and the trigger effect are assumed to be 0.

@@

# simulate toy data set for one subject 
d_init <-
  # create the 8 conditions
  tibble(prior = rep(c(0, 1), each = 4),
         comp = rep(c(0, 0, 1, 1), times = 2),
         rel = rep(c(0,1), times = 4),
         main_type = as.factor(sample(rep(c("xor", "some"), times=4)))) %>% 
  rowwise() %>%
  # sample effects based on condition
  mutate(
    subj_ID = 1, 
    intercept = 0,
    prior_obs = ifelse(prior == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, -0.15, 0.05)),
    comp_obs = ifelse(comp == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, 0.15, 0.05)),
    rel_obs = ifelse(rel == 0, rnorm(1, sd = 0.05), 
                       rnorm(1, 0.15, 0.05)),
    # sample response
    # what do we do about the interactions? can they safely be ignored?
    target = rnorm(1, 
                   mean = (intercept + prior_obs + comp_obs + rel_obs), 
                   sd = 1) # how to set / determine this properly?
                           # MF: irrelvant; we should z-score anyway exactly like we 
                           #     do for the real analysis 
                           #     this also means that choices of mean and SD for 
                           #     distributions to sample coefficents for are no that relevant
  )
d_init
```

@@

# MF's crowbar ignoramus approach

```{r, eval = FALSE}
# create fake data
create_fake_data <- function(N = 100, seed = 1) {
  set.seed(seed)
  beta_rel  <- 0.15
  beta_comp <- 0.15
  beta_pri  <- 0.15
  map_df(
    1:N, 
    function(i) {
      tibble(
        subj    = rep(str_c("subj_",i), times = 8),
        trigger = rep(c('or', 'some'), times= 4), 
        rel     = rnorm(8),
        comp    = rnorm(8),
        pri     = rnorm(8)
      ) %>% 
        group_by(subj) %>% 
        mutate(
          rel     = rel-mean(rel) / sd(rel),
          comp    = comp-mean(comp) / sd(comp),
          pri     = pri-mean(pri) / sd(pri)
        ) %>% 
        ungroup() %>% 
        mutate(
          target  = rnorm(8, beta_rel * rel + beta_comp * comp + beta_pri * pri)
        )
    }
  )
}
initial_fake_data <- create_fake_data(N=100, seed = 3)

# initial fit
fit_initial <- brm(target ~ trigger * rel * comp * pri, initial_fake_data)

# main results
N <- 300
k <-100

results <- map_dbl(
  1:k, 
  function(i) {
    fake_data <- create_fake_data(N=N, seed = i)
    fit <- update(fit_initial, newdata = fake_data, seed = i, cores = 4)
    lower_bounds <- rbind(
      aida::summarize_sample_vector(tidybayes::tidy_draws(fit) %>% select(b_rel), "rel"),
      aida::summarize_sample_vector(tidybayes::tidy_draws(fit) %>% select(b_comp), "comp"),
      aida::summarize_sample_vector(tidybayes::tidy_draws(fit) %>% select(b_pri), "pri")
    ) %>% pull("|95%")
    min(lower_bounds)
  }
)

message("Effective power for N=", N, " is:" , mean(results > 0.05))
